### 2018-07-22 fe_px_paper/fe_body_text.tex
Finished 1st draft of manuscript. Still finishing data analysis, but the language is good to go!

Manganese radionuclides are desirable for radiopharmaceutical applications, as manganese is a chemically versatile element, biologically relevant, and essential in trace quantities. 
Manganese possesses well-established biochemistry, and has been proven to be chelated well by DOTA, with high biostability at neutral pH \cite{Graves2015}.

paraphrase more of graves talk on mn applicatiosn here...

manganese is chelated well by DOTA , and has been shown to image antobodies for more than a week ofater administration

longer useful imaging time than Zr89 PET

While these efforts have been targeted towards the production cross section of the \ce{^{51,52g,52m}Mn} PET isotopes (as well as other emerging medical radionuclides), these measurements offer insight into the spin distribution of excited nuclear states, in the 5--55 MeV range.

% In addition to the interest in the production of 51,52Mn for PET research, this experiment offered an opportunity to study the distribution of angular momentum in compound nuclear and direct pre-equilibrium reactions via observation of the \ce{^{52g}Mn} (t½ =5.591±0.003 d; Jπ =6+) to \ce{^{52m}Mn} (t½ =21.1±0.2 m; Jπ =2+) ratio [5]. A set of measured cross sections for the various natFe(p,x), natTi(p,x), and natCu(p,x) reactions between 5–55 MeV, as well as independent measurements of isomer branching ratios, are reported.
% Figure 3 shows the 52mMn(2+)/52gMn(6+) isomer-to-ground state branching ratio for the naFe(p,αn) reaction, calculated using TALYS-v1.8 over the 20 - 55 MeV range spanned by the foil stack. It is possible to use this measurement to infer the spin cut-off σ2, the width of the angular momentum distribution of the level density [6], [7]. The isomer-to-ground state ratio lends insight into equilibrium / pre-equilibrium nature of reactions in this energy region, useful when attempting to minimize dose from an unwanted isomer/ground state of a medical radionuclide.

These measurements provide a range of cross section data invaluable to not only the medical isotope production community, but also as new measurements to improve the reliability of the range of modern reaction modeling codes. In addition, these experiments provide valuable insight into the challenges and unexpected nuances involved in precision cross section data measurements.

@@

@@

\begin{figure}
 \centering
 \includegraphics[width=0.5\linewidth]{../nb_px_paper/figures/86Y.pdf}
 % Nb_ptallies.pdf: 380x298 pixel, 72dpi, 13.41x10.51 cm, bb=0 0 380 298
 \caption{Measured \ce{^{93}Nb}(p,x)\ce{^{86}Y} cross section, with the \ce{^{93}Nb}(p,$\alpha$p3n)\ce{^{86}Y} reaction channel visibly peaking at approximately \mbox{70 MeV}.}
 \label{fig:86Y}
\end{figure}

>>>>

Manganese radionuclides are desirable for radiopharmaceutical applications, as manganese is a chemically versatile element, biologically relevant, and essential in trace quantities. 
Manganese possesses well-established biochemistry, and has been proven to be chelated well by DOTA for tracking monoclonal antibodies, with high biostability at neutral pH \cite{Graves2015}.
\ce{^{52}Mn} ($t_{1/2}$ = 5.591$\pm$0.003\,d, $I_{\beta^+}$ = 29.4\%, $E_{\beta\text{, avg}}$ = 0.242\,MeV) has been shown to be useful for immunoPET applications, with its rapid blood clearance offering the possibility for imaging within minutes of injection \cite{Dong2015}.
Its lifetime allows for immunoPET imaging up to 2--3 weeks after administration, giving \ce{^{52}Mn} a longer duration for useful imaging  than the more established immunoPET agents \ce{^{89}Zr} or   \ce{^{64}Cu} \cite{Graves2015}. 
\ce{^{52}Mn}  has high uptake in the heart, liver, kidneys, and pancreas, making it a useful diagnostic agent for pancreatic $\beta$-cell, insulinoma, and porphysome imaging. 
However, its  long lifetime and unfavorable decay gamma-rays (notably, its  744.233\,keV [$I_\gamma = 90.0\,\pm\,1.2\%$], 935.544\,keV [$I_\gamma = 94.5\,\pm\,1.3\%$], and 1434.092\,keV [$I_\gamma = 100.0 \pm 1.4\%$] lines) make it undesirable for clinical applications, but easily shipped worldwide and highly suitable for pre-clinical imaging \cite{Dong2015}.
The \ce{^{52m}Mn} isomer ($t_{1/2}$ = 21.1$\pm$0.2\,min, $I_{\beta^+}$ = 96.6\%, $E_{\beta\text{, avg}}$ = 1.172\,MeV) offers a much stronger positron emission branch than \ce{^{52g}Mn}, but its short lifetime makes production and handling difficult, and its  high-intensity gamma emission (1434.06\,keV [$I_\gamma = 98.2\,\pm\,0.5\%$]) contributes significantly to patient dose  \cite{Dong2015}.
Another emerging manganese radionuclide is \ce{^{51}Mn} ($t_{1/2}$ = 46.2$\pm$0.1\,min, $I_{\beta^+}$ = 97.09\%, $E_{\beta\text{, avg}}$ = 0.973\,MeV), whose short lifetime makes it clinically suited for rapid metabolic studies \cite{Wang2017}.
\ce{^{51}Mn} has a comparable positron branch and energy to \ce{^{52m}Mn}, more suitable for clinical use than the soft positron emissions of \ce{^{52g}Mn}. 
In addition, \ce{^{51}Mn} lacks any strong decay gamma-rays (its longer-lived daughter \ce{^{51}Cr} [$t_{1/2}$ = 27.704$\pm$0.003\,d] has only a single  320.0284\,keV [$I_\gamma = 9.910\,\pm\,0.010\%$] line), making it the best choice of these radionuclides for clinical imaging applications \cite{Wang2017}.

% While these efforts have been targeted towards the production cross section of the \ce{^{51,52g,52m}Mn} PET isotopes (as well as other emerging medical radionuclides produced in these irradiations), these measurements offer insight into the spin distribution of excited nuclear states, in the 5--55 MeV range.

% In addition to the interest in the production of\ce{^{51,52g,52m}Mn} for PET research

While these efforts have been motivated by the production cross section of the \ce{^{51,52g,52m}Mn} radionuclides for PET studies, this experiment offers an opportunity to study the distribution of angular momentum in compound nuclear and direct pre-equilibrium reactions via observation of the \ce{^{52m}Mn} ($t_{1/2}$ = 21.1$\pm$0.2\,min; J$^\pi=2+$) to \ce{^{52g}Mn} ($t_{1/2}$ = 5.591$\pm$0.003\,d; J$^\pi=6+$)   ratio \cite{Dong2015,Wang2017}.
Measurements of isomer-to-ground state ratios have been used for over 20\,years to probe the spin distribution of excited nuclear states in the A\,$\approx$\,190 region \cite{PhysRevC.73.034613,PhysRevC.45.1171}.
These measurements provide a range of cross section data invaluable to not only the medical isotope production community, but  also serve to measure the spin distribution in f$_{7/2}$ subshell nuclei via the observation of isomer-to-ground state ratios, and  as new data to improve the predictive capabilities of  reaction modeling codes. 
In addition, these experiments provide valuable insight into the challenges  involved in precision cross section data measurements.

% A set of measured cross sections for the various natFe(p,x), natTi(p,x), and natCu(p,x) reactions between 5–55 MeV, as well as independent measurements of isomer branching ratios, are reported.
% Figure 3 shows the 52mMn(2+)/52gMn(6+) isomer-to-ground state branching ratio for the naFe(p,αn) reaction, calculated using TALYS-v1.8 over the 20 - 55 MeV range spanned by the foil stack. It is possible to use this measurement to infer the spin cut-off σ2, the width of the angular momentum distribution of the level density [6], [7]. The isomer-to-ground state ratio lends insight into equilibrium / pre-equilibrium nature of reactions in this energy region, useful when attempting to minimize dose from an unwanted isomer/ground state of a medical radionuclide.

% While this might seem obvious, the contributions to the slowing of the beam due to the adhesive has often been neglected in much work performed to date. While plays a limited role at high beam energies, it becomes increasingly important for proton energies below 25 MeV.

% In addition to providing a potentially highly-valuable beam monitor, the Nb(p,x) reactions offer an opportunity to study the angular momentum deposition via pre-equilibrium reactions and the spin distribution in f$_{7/2}$ subshell nuclei via the observation of isomer-to-ground state ratios.

% lee stuff:
% 
% In addition to the interest in the production of 51,52 Mn for PET research, this experiment offered
% an opportunity to study the distribution of angular momentum in compound nuclear and direct
% pre-equilibrium reactions via observation of the 52g Mn (t ½ =5.591±0.003 d; J π =6 + ) to 52m Mn
% (t ½ =21.1±0.2 m; J π =2 + ) ratio. These results indicated were presented at the University of Oslo in
% May 2017 5 and indicate that the constant
% temperature nuclear level density model
% which has had the greatest success at
% reproducing most low-energy nuclear
% reaction data fails to reproduce the
% deposition of angular momentum at both
% compound and pre-compound energies
% (see figure 3). This result will be
% discussed in greater detail in Mr. Andrew
% Voyles’ doctoral dissertation, which is
% expected to be completed by May 2018.
% 
% An important outcome from this work has an appreciation for the role played by the latex
% adhesive on the Kapton tape used to contain the individual stacked targets. While this might
% seem obvious, the contributions to the slowing of the beam due to the adhesive has been
% neglected in work performed at LANL-IPF to date. While this is expected to play a limited role
% at high beam energies, it becomes increasingly important for proton energies below 25 MeV.

@@

% lee:
% 
% In addition to the interest in the production of 51,52 Mn for PET research, this experiment offered
% an opportunity to study the distribution of angular momentum in compound nuclear and direct
% pre-equilibrium reactions via observation of the 52g Mn (t ½ =5.591±0.003 d; J π =6 + ) to 52m Mn
% (t ½ =21.1±0.2 m; J π =2 + ) ratio. These results indicated were presented at the University of Oslo in
% May 2017 5 and indicate that the constant
% temperature nuclear level density model
% which has had the greatest success at
% reproducing most low-energy nuclear
% reaction data fails to reproduce the
% deposition of angular momentum at both
% compound and pre-compound energies
% (see figure 3). This result will be
% discussed in greater detail in Mr. Andrew
% Voyles’ doctoral dissertation, which is
% expected to be completed by May 2018.

@@

\begin{figure}
 \centering
 \includegraphics[width=0.5\linewidth]{./figures/51Mn_px.pdf}
 % Nb_ptallies.pdf: 380x298 pixel, 72dpi, 13.41x10.51 cm, bb=0 0 380 298
 \caption{Measured \ce{^{93}Nb}(p,x)\ce{^{86}Y} cross section, with the \ce{^{93}Nb}(p,$\alpha$p3n)\ce{^{86}Y} reaction channel visibly peaking at approximately \mbox{70 MeV}.}
 \label{fig:86Y}
\end{figure}
